---
title: "teal.code R Package Development Guide"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package Overview

`teal.code` is an R package for storing code and its execution environment in objects called `qenv`.
It supports reproducible code execution in teal Shiny applications. Developers can inspect and modify `qenv` objects,
providing full control over the stored code and environment. These objects also capture warnings and errors automatically.

## Development Context

The teal framework uses Shiny to create reproducible environments for data analysis. Within this framework, `teal.code`
manages code and its execution environment. Related packages include:

- `teal.reporter`: Integrates reports with reproducible code in teal applications. Its reporter functions accept `qenv`
    objects as inputs.
- `teal`: Provides the core architecture for teal Shiny applications. Teal modules should use `qenv` objects to ensure
    that code changes are recorded and reproducible.

For an introduction to `qenv` and its related functions and methods, see @vignettes/qenv.Rmd.

### Workflows

- When creating a `qenv`, favor readability over concise code.
- For simple `qenv` objects, favor `within()`, as it supports more readable code.
- When creating a `qenv` that uses functions with side effects, such as `runif()`, always use `eval_code()`.
- For a complex `qenv`, consider dividing the work into smaller `qenv` objects that encapsulate distinct logical steps.
- For a long or complex `qenv`, choose between `eval_code()` and `within()` based on:
        - the number of external variables added to the `qenv`; a larger number favors `eval_code()`;
        - how often the same external variable is used; greater repetition favors `within()`.
- When fixing bugs or adding features, avoid changing the core principles of `qenv`. Check whether an issue already
    exists. If the expected behavior is unclear, request that an issue be created before making changes that could have
    unexpected consequences.

```{r teal_content, echo=FALSE, results='asis'}
# Append common shared instructions from teal framework
common_claude_url <- "https://raw.githubusercontent.com/insightsengineering/teal/refs/heads/add_common_claude%40main/SHARED_CLAUDE.md" # nolint: line_length_linter.
common_claude_path <- tempfile()
download.file(url = common_claude_url, destfile = common_claude_path, quiet = TRUE)
common_claude_shared <- readLines(common_claude_path)
# remove CI gitlab content and github section comments
gitlab_starts <- grep("gitlab starts here", common_claude_shared)
gitlab_ends <- grep("gitlab ends here", common_claude_shared)
common_claude_shared <- common_claude_shared[-1*seq(gitlab_starts, gitlab_ends)]
github_starts <- grep("github starts here", common_claude_shared)
github_ends <- grep("github ends here", common_claude_shared)
common_claude_shared <- common_claude_shared[-1*c(github_starts, github_ends)]
cat(common_claude_shared, sep = "\n")
```