---
output: github_document
---
# teal.code R Package Development Guide

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r render, eval=FALSE, include=FALSE}
rmd_path <- "vignettes/_AGENTS.Rmd"

rmarkdown::render(
  input = rmd_path,
  output_format = "github_document",
  output_file = "AGENTS.md",
  output_dir = ".",
  quiet = TRUE
)

file.remove("AGENTS.html")
```

## Package Overview

`teal.code` is an R package for storing code and its execution environment in objects called `qenv`.
It supports reproducible code execution in teal Shiny applications. Developers can inspect and modify `qenv` objects,
providing full control over the stored code and environment. These objects also capture warnings and errors automatically.

## Development Context

The teal framework uses Shiny to create reproducible environments for data analysis. Within this framework, `teal.code`
manages code and its execution environment. Related packages include:

- `teal.reporter`: Integrates reports with reproducible code in teal applications. Its reporter functions accept `qenv`
    objects as inputs.
- `teal`: Provides the core architecture for teal Shiny applications. Teal modules should use `qenv` objects to ensure
    that code changes are recorded and reproducible.

For an introduction to `qenv` and its related functions and methods, see @vignettes/qenv.Rmd.

### Workflows

- When creating a `qenv`, favor readability over concise code.
- For simple `qenv` objects, favor `within()`, as it supports more readable code.
- When creating a `qenv` that uses functions with side effects, such as `runif()`, `library` or `options` always use `eval_code()`. Use the appropriate comment to link side effects to objects
- For a complex `qenv`, consider dividing the work into smaller `qenv` objects that encapsulate distinct logical steps.
- For a long or complex `qenv`, choose between `eval_code()` and `within()` based on:
    - the number of external variables added to the `qenv`; a larger number favors `eval_code()`;
    - how often the same external variable is used; greater repetition favors `within()`.
- When creating custom functions within a teal app server, it is better to do define them inside a qenv. If defined outside
the function might fail.
- When fixing bugs or adding features, avoid changing the core principles of `qenv`. Check whether an issue already exists. If the expected behavior is unclear, request that an issue be created before making changes that could have
unexpected consequences.

This package is part of the teal framework. The following configuration applies to all packages within the teal framework:

```{r teal_content, echo=FALSE, results='asis'}
# Append common shared instructions from teal framework
common_agents_url <- "https://raw.githubusercontent.com/insightsengineering/teal/refs/heads/main/AGENTS.md" # nolint: line_length_linter.
common_agents_path <- tempfile()
download.file(url = common_agents_url, destfile = common_agents_path, quiet = TRUE)
common_agents_shared <- readLines(common_agents_path)
# remove a heading line only intented to aovid a markdown lint error
line_to_avoid_lintr <- grep("^# ", common_agents_shared)
common_agents_shared <- common_agents_shared[-1 * line_to_avoid_lintr]
cat(common_agents_shared, sep = "\n")
```
