---
# This file generates AGENTS.md in the package root.
# It is not part of the package build or the pkgdown site.
# To regenerate it, run from the package root:
# rmarkdown::render("vignettes/_AGENTS.Rmd", output_file = file.path("AGENTS.md"), output_dir = ".")
output:
  github_document:
    html_preview: false
params:
  url: "https://raw.githubusercontent.com/insightsengineering/teal/refs/heads/main/SHARED_AGENTS.md"
---
# teal.code R Package Development Guide

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package Overview

`teal.code` is an R package for storing code and its execution environment in objects called `qenv`.
It supports reproducible code execution in teal Shiny applications. Developers can inspect and modify `qenv` objects,
providing full control over the stored code and environment. These objects also capture warnings and errors automatically.

## Development Context  

The teal framework uses Shiny to create reproducible environments for data analysis. Within this framework, `teal.code`
manages code and its execution environment. Related packages include:

- `teal.data`: Stores all data objects, data relations and
  transformations of data used in teal applications, in an object called
  `teal_data`. The `teal_data` object inherits from the `qenv` and therefore 
  shares all properties about code reproducibility.
- `teal`: Provides the core architecture for teal Shiny applications.
  Teal modules should use `teal_data` objects to ensure that code changes are
  recorded and reproducible.
- `teal.reporter`: Integrates reports with reproducible code in teal
  applications. Its reporter functions accept `teal_data` objects as inputs and
  extend this class.

For an introduction to `qenv` and its related functions and methods, see @vignettes/qenv.Rmd.

### Workflows

- When creating a `qenv`, favor readability over concise code.
- For simple `qenv` objects, favor `within()`, as it supports more readable code.
- When creating a `qenv` that uses functions with side effects, such as `runif()`, `library` or `options` always use `eval_code()`. Use the appropriate comment to link side effects to objects
- For a complex `qenv`, consider dividing the work into smaller `qenv` objects that encapsulate distinct logical steps.
- In the line of the previous comment, `qenv` can be combined with method `c.qenv`. Be careful that they cannot have elements with
the same name and about the combined order of code execution.
- For a long or complex `qenv`, choose between `eval_code()` and `within()` based on:
    - the number of external variables added to the `qenv`; a larger number favors `eval_code()`;
    - how often the same external variable is used; greater repetition favors `within()`.
- When creating custom functions within a teal app server, it is better to do define them inside a qenv. If defined outside
the function might fail.
- When fixing bugs or adding features, avoid changing the core principles of `qenv`. Check whether an issue already exists. If the expected behavior is unclear, request that an issue be created before making changes that could have
unexpected consequences.

### Debugging `qenv`

Errors in teal applications commonly occur while a `qenv` is being evaluated. By default, `qenv` handles these errors
and returns an object of class `qenv.error`. This prevents the teal application from crashing immediately and provides
the user with a helpful error message.

After identifying the problematic `qenv`, use the following steps to investigate it:

- Use the `qenv.error` message to identify the line of code that triggered the error.
- If that line alone does not explain the error, inspect the preceding code. If necessary, create a temporary version
  of the `qenv` containing only the preceding calls, or modify the line that triggered the error.
- Inspect values stored in the `qenv` using `qenv[["my_value"]]` to help identify the source of the error.
- Because messages and warnings produced by a `qenv` are not printed to the R console, inspect them using
  `get_messages()` and `get_warnings()`.
- The same techniques for inspecting values, messages, and warnings can be used when a `qenv` produces incorrect
  results without raising an error.

```{r teal_content, echo=FALSE, results='asis'}
# Append common shared instructions from teal framework
common_agents_url <- params$url
common_agents_path <- tempfile()
download.file(url = common_agents_url, destfile = common_agents_path, quiet = TRUE)
common_agents_shared <- readLines(common_agents_path)
# remove optional sections
remove_sections <- c("<!-- (Begins|Ends) Module Section")
Reduce(
  function(x, y) {
    delim <- grep(y, x)
    if (length(delim) == 0) {
      return(x)
    } else if (length(delim) != 2) {
      stop("Section delimiters are not balanced")
    }
    x[-1 * seq(from = delim[1], to = delim[2])]
  },
  remove_sections,
  init = common_agents_shared
) |>
  cat(sep = "\n")
```
