---
# This file generates AGENTS.md in the package root.
# It is not part of the package build or the pkgdown site.
# To regenerate it, run from the package root:
# rmarkdown::render("vignettes/_AGENTS.Rmd", output_file = file.path("AGENTS.md"), output_dir = ".")
output:
  github_document:
    html_preview: false
params:
  url: "https://raw.githubusercontent.com/insightsengineering/teal/refs/heads/main/SHARED_AGENTS.md"
---
# teal.code R Package Development Guide

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package Overview

`teal.code` is an R package for storing code and its execution environment in objects called `qenv`.
It supports reproducible code execution in teal Shiny applications. Developers can inspect and modify `qenv` objects,
providing full control over the stored code and environment. These objects also capture warnings and errors automatically.

## Development Context

The teal framework uses Shiny to create reproducible environments for data analysis. Within this framework, `teal.code`
manages code and its execution environment. Related packages include:

- `teal.reporter`: Integrates reports with reproducible code in teal applications. Its reporter functions accept `qenv`
objects as inputs.
- `teal`: Provides the core architecture for teal Shiny applications. Teal modules should use `qenv` objects to ensure
that code changes are recorded and reproducible.
- `teal.data`: Stores all data objects, data relations and transformations of data used in teal applications. The `teal.data`
object is based on the `qenv` and therefore used for code reproducibility.

For an introduction to `qenv` and its related functions and methods, see @vignettes/qenv.Rmd.

### Workflows

- When creating a `qenv`, favor readability over concise code.
- For simple `qenv` objects, favor `within()`, as it supports more readable code.
- When creating a `qenv` that uses functions with side effects, such as `runif()`, `library` or `options` always use `eval_code()`. Use the appropriate comment to link side effects to objects
- For a complex `qenv`, consider dividing the work into smaller `qenv` objects that encapsulate distinct logical steps.
- In the line of the previous comment, `qenv` can be combined with method `c.qenv`. Be careful that they cannot have elements with
the same name and about the combined order of code execution.
- For a long or complex `qenv`, choose between `eval_code()` and `within()` based on:
    - the number of external variables added to the `qenv`; a larger number favors `eval_code()`;
    - how often the same external variable is used; greater repetition favors `within()`.
- When creating custom functions within a teal app server, it is better to do define them inside a qenv. If defined outside
the function might fail.
- When fixing bugs or adding features, avoid changing the core principles of `qenv`. Check whether an issue already exists. If the expected behavior is unclear, request that an issue be created before making changes that could have
unexpected consequences.

### Debugging `qenv`

When debugging teal apps, it is very common that the error happens within the execution of a `qenv`. It is important
to know how to debug `qenv`.
Firstly, by default when there is an error within a qenv its automatically handled, and a
`qenv.error` class object will be returned. That will prevent the teal app to immediately crash and will display
an useful error message to the user.
Once identified a problematic `qenv`, the guidelines to debug it are:
- As the `qenv` does not print messages and warnings to the R console, so explore them with functions `get_messages` and `get_warnings`.
- Access individual elements of the `qenv` with `get_var[[qenv, "my_var"]]` to inspect internal values.
- Use function `get_code` to extract specifically how an element is created. That will help in identifying errors.

```{r teal_content, echo=FALSE, results='asis'}
# Append common shared instructions from teal framework
common_agents_url <- params$url
common_agents_path <- tempfile()
download.file(url = common_agents_url, destfile = common_agents_path, quiet = TRUE)
common_agents_shared <- readLines(common_agents_path)
# remove a heading line only intented to aovid a markdown lint error
line_to_avoid_lintr <- grep("^# ", common_agents_shared)
common_agents_shared <- common_agents_shared[-1 * line_to_avoid_lintr]
cat(common_agents_shared, sep = "\n")
```
